---
title: "R Seminar Week 4 -- For Loops"
author: "John Kearns"
date: "2023-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# this is the initial set up
# include = FALSE tells R to not show this chunk of code when it knits the page
# echo = TRUE sets it as a default to show the output of code in the final document

library(tidyverse)
library(readxl)

```

```{r,echo=FALSE,include=FALSE}

#master_dir
# fill in your working directory here

```


How does a for-loop work
```{r}

for(i in 1:5){
  
  print(i)
  
}

```

If and else statements

```{r}

x = 2

if(x%%2==0){
  
  print("This is even")
  
} else{
  
  print("This is odd")
  
}


```
Combining for and if statements in a user-defined function
```{r}

even_or_odd = function(numbers){
  
  for(num in numbers){
    
    if(num%%2==0){
  
  print("This is even")
  
} else{
  
  print("This is odd")
  
}
    
  }
  
}

```

```{r,error=TRUE}

even_or_odd(1)

even_or_odd(c(1,3,4,5,6,7))

even_or_odd(NA)

```

What if we wanted to read in a bunch of sheets from an excel file?

Look at the food dataset from the Census Bureau. Starting with the first sheet, what would we have to do to read this in a way we could work with the data?

```{r}

data = read_excel("C:/Users/16094/Documents/Year 2 - Princeton/Fall 2023/R Tutoring/SPIA_R_Seminars/Week 4/Data/Raw/food1_week62.xlsx")
?read_excel()

```

```{r}

data = read_excel("C:/Users/16094/Documents/Year 2 - Princeton/Fall 2023/R Tutoring/SPIA_R_Seminars/Week 4/Data/Raw/food1_week62.xlsx",range="A8:G199",col_names=FALSE)

colnames(data) = c("characteristic","total","enough_wanted","enough_not_wanted","sometimes_not_enough","often_not_enough","not_report")

# pick a characteristic we want to visualize


```

How could we use a for-loop to make our life easier?

```{r,warning=FALSE}

state_df = data.frame()

for(state in c("US",state.abb)){
  


}


```

Now make a pdf with a bar chart for each state

```{r}
pdf(file="C:/Users/16094/Documents/Year 2 - Princeton/Fall 2023/R Tutoring/SPIA_R_Seminars/Week 4/Charts/state_bar_charts.pdf",height=10,width=10)
for(stat1 in unique(state_df$state)){

plot1 = 

print(plot1)

}
dev.off()

```

