---
title: "R Seminar Week 4 -- For Loops"
author: "John Kearns"
date: "2023-11-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# this is the initial set up
# include = FALSE tells R to not show this chunk of code when it knits the page
# echo = TRUE sets it as a default to show the output of code in the final document

library(tidyverse)
library(readxl)

```

```{r,echo=FALSE,include=FALSE}

#master_dir
# fill in your working directory here

```

General context: We are trying to visualize employment data for some areas in NJ for the years 2016 to 2023

### 1. Pull the data for the following area codes from the Bureau of Labor Statistics API, using a loop

The following links will be helpful: 
"https://data.bls.gov/cew/doc/titles/industry/industry_titles.csv"
"https://data.bls.gov/cew/doc/titles/area/area_titles.csv"
"https://www.bls.gov/cew/classifications/ownerships/sic-ownership-titles-csv.csv"
"https://data.bls.gov/cew/data/api/2023/1/area/CS408.csv" # what is "2023/1/" referring to?


```{r}
codes = c("US000",34001,34003,34005,34007,34009,34011,34013,34015,34017,34019,34021
,34023,34025,34027,34029,34031,34033,34035,34037,34039,34041)

```

```{r,echo=FALSE}

# code to read in data for each of the codes
# recommendation: make sure the loop works for one of the codes, before running the loop for all of them



```

### 2. In a new dataset, keep data that is ONLY the total covered employment for all industries. Join the information on for industry titles and area titles

```{r}



```

### 3. Visualize average weekly wage for the US and each county in NJ on one line plot. Change the color theme of the plot and add some labels. *If time*, make Mercer County a princeton orange color, and the rest of the counties/areas black

```{r}




```
### 4. In a new dataset, pull the following data for Mercer County: 2023 data, all industry codes less than 3 characters long and at least 100 employed persons in month 3. Make a bar chart showing the location quotients for the 5 highest and lowest industries in terms of location quotient. A location quotient is a measure of how much of the population is employed in a specific industry relative to the general population.

```{r}



```


### 5. Using the all-county employment dataset, plot the change in average weekly wage in each country relative to the first year available. Save an individual plot for each county in a new pdf file. Which counties have grown the fastest and slowest?


```{r}



```

