---
title: "R Seminar Week 3 -- Visualizing Data"
author: "John Kearns"
date: "2023-10-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# this is the initial set up
# include = FALSE tells R to not show this chunk of code when it knits the page
# echo = TRUE sets it as a default to show the output of code in the final document

library(tidyverse)
library(usmap)

```

```{r,echo=FALSE,include=FALSE}

master_dir = "C:/Users/16094/Documents/Year 2 - Princeton/Fall 2023/R Tutoring/SPIA_R_Seminars/Week 3/"
do_folder = paste0(master_dir,"Do/")
data_folder = paste0(master_dir,"Data/")
results_folder = paste0(master_dir,"Results/")
charts_folder = paste0(master_dir,"Charts/")

```


Let's load in the data we are working with today
```{r}

payment_data = read_csv(paste0(data_folder,"Raw/payment_data.csv"))

```
## Data transformations

Get state names in the data set

Hint: the numbers refer to the alphabetized state abbreviations + 51 for DC

```{r}





```


Indicator for if they have ever used Apple, Google, or Samsung Pay
```{r}

payment_data = payment_data %>% 
  mutate(mobile_apps = ifelse(apple_pay_adopt==1|google_pay_adopt==1|samsung_pay_adopt==1,1,0))

```


## Let's make a chart showing rate of app adoption by age

First make the dataset
```{r}

chart1_data = payment_data %>% 
  group_by(age) %>% 
  summarize(mobile_apps = mean(mobile_apps,na.rm=TRUE)) %>% 
  ungroup()

```
 
 
Now make the chart 
```{r}

ggplot(chart1_data,aes(x=age,y=mobile_apps)) +
  geom_point()

```

That is a little ugly...
Graphic options: themes, labels, sizes, colors

```{r}

ggplot(chart1_data,aes(x=age,y=mobile_apps)) +
  geom_line(size=2)

ggplot(chart1_data,aes(x=age,y=mobile_apps)) +
  geom_line(size=2,color="darkorange")

ggplot(chart1_data,aes(x=age,y=mobile_apps)) +
  geom_line(size=2,color="darkorange") +
  theme_bw()

ggplot(chart1_data,aes(x=age,y=mobile_apps)) +
  geom_line(size=2,color="darkorange") +
  theme_bw() +
  labs(x="Age",y="Adoption Rate",title="Adoption rate of Apple, Google, or Samsung Pay",caption="Source: Fed Atlanta")

```


# How about visualizing access to savings account by race 

```{r}

chart2_data = payment_data %>% 
  mutate(race_txt = case_when(
    hispaniclatino==1~"Hispanic",
    race==1~"White",
    race==2~"Black",
    race==3~"Native American",
    race==4~"Asian",
    race==5~"Hawaiian or Pacific Isl.",
    race==6~"Mixed Ethnicity"
  )) %>% 
  group_by(race_txt) %>% 
  summarize(bank_acct = mean(have_savings_acct,na.rm=TRUE))

```

Make the chart
```{r}

ggplot(chart2_data,aes(x=race_txt,y=bank_acct,fill=race_txt)) +
  geom_bar(stat="identity")

# what do we want to change? Get rid of NA, reorder bars, anything else?


```


# Looking at relationship between income and buying things on computer by race?
```{r}

chart3_data = payment_data %>% 
  mutate(race_txt = case_when(
    hispaniclatino==1~"Hispanic",
    race==1~"White",
    race==2~"Black",
    race==3~"Native American",
    race==4~"Asian",
    race==5~"Hawaiian or Pacific Isl.",
    race==6~"Mixed Ethnicity"
  )) %>% 
  group_by(hhincome,race_txt) %>%
  summarize(computer_use=mean(computer_adopt,na.rm=TRUE))


```
```{r}

ggplot(chart3_data,aes(x=hhincome,y=computer_use,color=race_txt)) +
  geom_line(size=2) +
  facet_wrap(~race_txt) +
  theme_bw()

```


Gender roles in financial situations?

Make the dataset
```{r}


```

```{r}




```



# Making quick maps

```{r}

chart5_data = payment_data %>% 
  group_by(state) %>% 
  summarize(mobile_instore = mean(mobile_inperson_adopt,na.rm=TRUE)) 

```

```{r}

chart1 = usmap::plot_usmap(data = chart5_data, values = "mobile_instore") + 
  scale_fill_continuous(name = "Mobile Payment Adoption Rate",type="viridis",trans="reverse") + 
  theme(legend.position = "right")

ggsave(paste0(charts_folder,"mobile_adoption.png"),chart1,width=10,height=10,unit="in")

```


