---
title: "R Seminar Week 2 -- Cleaning Data"
author: "John Kearns"
date: "2023-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# this is th initial set up
# include = FALSE tells R to not show this chunk of code when it knits the page
# echo = TRUE sets it as a default to show the output of code in the final document

library(tidyverse)

```

## RMarkdown basics

A chunk is where R code is written. You can add a new chunk with ctrl+alt+i

adding include=FALSE or echo=FALSE will affect how the code is shown in the final output

knitting refers to running all of the code and saving it as a certain output (html,pdf,latex)

Headers can be made with #s
Largest header is one # and lowest is six #s

You can write equations and do fancy things pretty easily in RMarkdown
Some more resources [here](https://sharp-lamport-f1a531.netlify.app/#1) 

## Let's look at a dataset


### Reviewing data frames

How do we make a data frame again? What is a data frame made up of?

```{r}



```




### Loading data in

Tip: make strings to refer to file structures so that other people can change it easily and run your code

```{r,echo=FALSE,include=FALSE}

master_dir = "C:/Users/16094/Documents/Year 2 - Princeton/Fall 2023/R Tutoring/SPIA_R_Seminars/Week 2/"
do_folder = paste0(master_dir,"Do/")
data_folder = paste0(master_dir,"Data/")
results_folder = paste0(master_dir,"Results/")
charts_folder = paste0(master_dir,"Charts/")

```


#### There are different ways to load data into R
You can use point-and-click method by going File -> Import Dataset -> select the type of data you are inserting

I recommend using the readr version when using csv files
If reading in a txt files or some other dataset that isn't comma-separated, use read.table() or read.delim()

```{r,echo=FALSE}

citibike = read.csv(paste0(data_folder,"Raw/citibike.csv"))

citibike = read.delim(paste0(data_folder,"Raw/citibike.txt"))

```


### Let's work with some messy data 
```{r}


covid = read_csv(paste0(data_folder,"Raw/covid_cases_deaths.csv"))


```

What do we notice about that data?


What types are the variables in the dataset?
```{r}

sapply(covid,class)

```


What changes do we think we need to do to the dataset?






## mutate

mutate() makes new variables and appends them to the end of the dataset

```{r}

covid1 = covid %>% 
  mutate(population=as.numeric(population))

```
What is the error telling us?


```{r}

covid1 = covid %>% 
  mutate(pop1 = gsub(",","",population),
         pop1 = as.numeric(pop1))

```


```{r}

covid2 = covid %>% 
  mutate(pop1 = gsub(",","",population)) %>% 
  mutate_at(vars(pop1),as.numeric)

identical(covid1,covid2)

```

## rename

rename() will allow you to more cleanly adjust column names

```{r}

covid1 = covid1 %>% 
  rename(diabetes_prevalence = `diabetes prevalence`)

```



## working with dates

```{r}

covid1 = covid1 %>% 
  mutate(date=as.Date(date,format="%m/%d/%Y"))

```

What can we do with a date variable?

```{r}
library(zoo)
library(lubridate)

month(covid1$date[1])
year(covid1$date[1])
quarter(covid1$date[1])
days_in_month(covid1$date[1])

```


## Calculations
What are some ways to do calculations?
To see summary statistics of a variable in a dataframe or a vector, use summary()
To add up values in a vector, use sum()
To get the mean, use mean()
To get the median, use median()
To get variance and standard deviation, use var() and sd()

max() and min() will give you maxmimum and minimum values, while which.max() would give you the index of the value that is the maximum in a vector

BEWARE OF NAs

To test or compare values, you can use inequalities. The code will give back a TRUE or FALSE.


```{r}

# lets start with getting a number of unique countries
# length() gives you how long a vector is
# unique() removes duplicates from a vector

length(covid1$location)
nrow(covid1)
ncol(covid1)
unique(covid1$location)
length(unique(covid1$location)) # 244 different countries


```

```{r}
# what are some statistics about covid cases?

summary(covid1)

summary(covid1$new_cases)

mean(covid1$new_cases) # why is this not giving a number back?

sd(covid1$new_cases)

```
## Grouping and summarizing

```{r}

# now let's get the average life expectancy by continent

# What are the values in the gender column?
unique(covid1$continent)

# what if we only want to see observations from Asia? How could we make that as a new dummy variable? 

# code for getting the data average life expectancy for Asian countries only:


# What would the steps be for getting average life expectancy for all continents at once:

covid1 %>% 
  group_by(continent) %>% # tells the code to apply whatever comes next to each group individually
  summarize(avg_life_expectancy = mean(life_expectancy,na.rm=TRUE)) %>% 
  ungroup() # in case you end up saving this as an object and doing other stuff, tell R to stop grouping


# What is the average age of the person in the dataset? 




```
#### What is this empty continent? Let's use a filter
```{r}

empty_continent = covid1 %>% 
  filter(continent=="")

# code if we want to remove these observations


```


#### Let's add another dataset
```{r}

osi = read_csv(paste0(data_folder,"Raw/osi.csv"))

```

How should we clean this up?

```{r}

osi1 = osi %>% 
  select(-c(1,2)) %>% # which variables do we want to drop?
  pivot_longer(cols="01Jan2020":"22May2022",values_to="osi",names_to="date") %>%
  mutate(date=as.Date(date,format="%d%b%Y"))

```


## merging data sets

```{r}

combined = left_join(covid2,osi1,by=c("iso_code"="country_code","date"="date"))

```


## Putting it all together

Let's make a dataset that gets (1) the average new cases per month in each country; (2) calculates the average monthly rate of growth; (3) spits out the average OSI and the average case growth rate by country

```{r}



```

Extra credit: slice this dataset so that it gives you only the country with the highest average OSI for each continent

```{r}


```


### Let's save our progress
```{r}

save.image(paste0(data_folder,"Processing/progress_20231013.RData"))

write.csv(combined,paste0(data_folder,"Final/covid-final.csv"))

```
