---
title: "Week 5 - Regressions and Tables"
author: "John Kearns"
date: "2023-11-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# installing the packages you need

#install.packages(c("lmtest","broom","estimatr","texreg","vtable","stargazer","robustbase"))

library(tidyverse)
library(robustbase)

```

### Load data in

```{r}

data("education")

education = education %>% 
    rename(residents = X1,
           per_capita_income = X2,
           young_residents = X3,
           per_capita_exp = Y,
           state = State) %>% 
    mutate(region = case_when(
        Region == 1 ~ "northeast",
        Region == 2 ~ "northcenter",
        Region == 3 ~ "south",
        Region == 4 ~ "west"
    )) %>% 
    select(-Region)

```


### summary statistics tables

```{r}
library(vtable)

st(education)

```
That is pretty ugly...but what is each part coming from?

```{r}

st(education %>% 
     select(residents,per_capita_income,per_capita_exp,region) %>% 
     rename("Number of residents"=residents,
            "State income per capita"=per_capita_income,
            "State per capita education expenditure"=per_capita_exp,
            "State region"=region),
   digits=2)


```

### t-tests

```{r}

t.test(education$per_capita_exp[education$region=="west"],education$per_capita_exp[education$region=="south"])

```

### regressions


### Make a plot

```{r}

ggplot(education,aes(x=per_capita_income,y=per_capita_exp,color=region)) +
  geom_point() +
  geom_smooth(method="lm",inherit.aes = FALSE,aes(x=per_capita_income,y=per_capita_exp)) +
  stat_poly_eq(use_label(c("eq", "R2")))

```



#### OLS

```{r}

lmfit = lm(per_capita_exp ~ residents, data = education)

summary(lmfit)

```

#### adding fixed effects

```{r}

lmfit = lm(per_capita_exp ~ residents + young_residents + per_capita_income + factor(region), data = education)

summary(lmfit)

```

### interaction terms

```{r}

education = education %>% 
  mutate(northeast = ifelse(region=="northeast",1,0))

lmfit = lm(per_capita_exp ~ per_capita_income*northeast, data = education)

```

### changing standard errors

To get heteroskedasticity-robust SEs, which should be your default.

Option 1:
```{r}

library(sandwich)
library(lmtest)

coeftest(lmfit, vcov = vcovHC(lmfit))

coeftest(lmfit, vcov = vcovHC(lmfit,type='HC1'))

```


Option 2:

```{r}
library(estimatr)

lmfit = lm_robust(per_capita_exp ~ residents + young_residents + per_capita_income + factor(region), data = education,se_type="stata")

summary(lmfit) # see how this is identical to the model we last ran

```
A drawback of lm_robust is that it is not optimized for working with time series standard errors/autocorrelation. The vcov() method might be better for this.


### Making regression tables

saving regression output as an object: Will only work with lm() objects

```{r}

library(broom)

lmfit = lm(per_capita_exp ~ residents + young_residents + per_capita_income + factor(region), data = education)

tidy_lm = tidy(lmfit)

tidy_lm

```

Exporting regression tables:

```{r}

library(texreg)

mod1 = lm(per_capita_exp ~ residents, data = education)
mod2 = lm(per_capita_exp ~ residents + young_residents, data = education)
mod3 = lm(per_capita_exp ~ residents + young_residents + per_capita_income, data = education)
mod4 = lm(per_capita_exp ~ residents + young_residents + per_capita_income + factor(region), data = education)

htmlreg(mget(c("mod1","mod2","mod3","mod4")), # mget() pulls objects with the given names
        include.ci=FALSE,  # shows standard errors instead of confidence intervals
        file="regression_results.html",
        caption="Regression results in order", # title of table
        #omit.coef=c(), put the names of any variables you do not want to show in the regression output
        #custom.gof.rows=list("name"=values), if you want to add another row of summary statistics 
        custom.model.names=c("Simple","Add young","Add income","Add region"),
        include.rsquared = TRUE,
        include.adjrs=FALSE,
        include.rmse=FALSE, 
        include.nobs = TRUE,
        start = c(.01,.05,.1)
        #reorder.gof=c(1,4,5,6,7,2,3) gives the order of the summary statistics rows
        )


```

