---
title: "R Seminar Week 1 -- Loading Data"
author: "John Kearns"
date: "2023-10-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# this is th initial set up
# include = FALSE tells R to not show this chunk of code when it knits the page
# echo = TRUE sets it as a default to show the output of code in the final document

library(tidyverse)

```

## RMarkdown basics

A chunk is where R code is written. You can add a new chunk with ctrl+alt+i

adding include=FALSE or echo=FALSE will affect how the code is shown in the final output

knitting refers to running all of the code and saving it as a certain output (html,pdf,latex)

Headers can be made with #s
Largest header is one # and lowest is six #s

You can write equations and do fancy things pretty easily in RMarkdown
Some more resources [here](https://sharp-lamport-f1a531.netlify.app/#1) 

## Let's look at a dataset

### Loading data in

Tip: make strings to refer to file structures so that other people can change it easily and run your code

```{r,echo=FALSE,include=FALSE}

master_dir = "C:/Users/16094/Documents/Year 2 - Princeton/Fall 2023/R Tutoring/SPIA_R_Seminars/Week 1/"
do_folder = paste0(master_dir,"Do/")
data_folder = paste0(master_dir,"Data/")
results_folder = paste0(master_dir,"Results/")
charts_folder = paste0(master_dir,"Charts/")

```


#### There are different ways to load data into R
You can use point-and-click method by going File -> Import Dataset -> select the type of data you are inserting

I recommend using the readr version when using csv files
If reading in a txt files or some other dataset that isn't comma-separated, use read.table() or read.delim()

```{r,echo=FALSE}

citibike = read_csv(paste0(data_folder,"Raw/citibike.csv"))
# what other things can you do with read_csv?

citibike = read.delim(paste0(data_folder,"Raw/citibike.txt"))

```

##### Enter the code here to import the excel version in our files. What about downloading staright from the web from this [link](https://github.com/amueller/introduction_to_ml_with_python/blob/master/data/citibike.csv)?

```{r}





```



#### What is in this dataset?
```{r}

head(citibike,30)

```

#### What are some things that we want to know about this data?
Number of men and women? What years does the data cover? Average length of ride?


What are some ways to do calculations?
To see summary statistics of a variable in a dataframe or a vector, use summary()
To add up values in a vector, use sum()
To get the mean, use mean()
To get the median, use median()
To get variance and standard deviation, use var() and sd()

max() and min() will give you maxmimum and minimum values, while which.max() would give you the index of the value that is the maximum in a vector

BEWARE OF NAs

To test or compare values, you can use inequalities. The code will give back a TRUE or FALSE.

```{r}

2<3

10>15

3==4

3!=4

c(1,2,3,4,5)<=3

"Hello"=="World"

"Hello"=="Hello"


```


```{r}

# lets start with getting a number of unique bikes
# length() gives you how long a vector is
# unique() removes duplicates from a vector

length(citibike$bikeid)
unique(citibike$birth.year)
length(unique(citibike$bikeid)) # 2,989 differnt bikes are used in this data set


```

```{r}
# what are some statistics about the length of rides?

summary(citibike)

summary(citibike$tripduration)

mean(citibike$tripduration)

sd(citibike$tripduration)

mean(citibike$birth.year) # why is this giving an error?


```

```{r}

# now let's get the number of rides by men and women

# What are the values in the gender column?
unique(citibike$gender)

# what if we only want to see rides by women? How could we make that as a new variable? 

# code for getting the data frame that only has rides by women:


# code for the number of rides by women? By men?


# What is the average age of the person in the dataset? 




```
#### What are the most common destinations?
```{r}


table(citibike$end.station.name)


# not easy to see what was the most common?
```
```{r}

sort(table(citibike$end.station.name),decreasing=TRUE)[1:10]

```

### Let's save our progress
```{r}

save.image(paste0(data_folder,"Processing/progress_20231006.RData"))

write.csv(paste0(data_folder,"Final/citibike-final.csv"))

```
#### FOR EXTRA TIME

Is there a difference in trip time at rush hour? Late at night?

Find the number of trips by the two types of users?

Are women or men more likely to be subscribers?

```{r}



```

